Adding selection to \stdpopsim: oh how the tables have turned

### 2025-01-08 analysis2_manuscript.tex
Update analysis2_manuscript.tex

    In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    In particular, variation in accuracy of the parameters estimated in each
    sample population depends on the methods used, and each method seems to 
    have its own biases with respect to the way in which population history 
    affects parameter estimates. For instance, for \dadicli we see that estimates of 
    the shape parameter are most accurate in the YRI sample, while for \polydfe
    estimates are least accurate in the YRI sample. 
    With respect to the strength of selection, as in constant size populations,
    all methods are performing reasonably well, although \polydfe seems to
    lead in terms of accuracy.
\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things.

@@

    In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    from the same three DFE-inference packages on simulations of the
    vaquita porpoise genome. In both situations, we see that all methods
    perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter. 
    A critical distinction is that the previous human simulations employed a DFE that assumed all selected mutations had additive effects
    (no dominance, $h = 1/2$), whereas the vaquita simulations use a DFE in which more deleterious mutations are more recessive ($h < 1/2$).
    For all three software packages, the inferred mean and shape of the distribution of selection coefficients 
    is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient
    and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
    Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
    so these results suggest caution in interpreting DFE inferences.

@@

    \subsection*{Simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. As stated above, simulations did or did not include a model 
    of (negative) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
    Three replicates of whole-genome simulations were performed for each species, with 
    a sample size of 100 diploid individuals. Demographic models are as stated above, each
    using the exon annotations from the human and vaquita porpoise genomes respectively.
    For simulations with selection, we sampled 10 diploid individuals from each of two populations (YRI and CEU). 
    To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations.

@@

    \subsection*{Distribution of fitness effects inference}
    We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
    \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
    Each software package was run using the same set of simulations.
    For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
    A key distinction between our analyses and those done in practice
    is that we have complete knowledge of which sites are under selection in the simulation,
    thus our inferences is a best case scenario. Empirical inference of the DFE is a much
    harder problem as some mutations assumed to be free of selection
    may in reality be under selection. As above, all methods were run using default settings.

>>>>

    In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    In particular, variation in accuracy of the parameters estimated in each
    sample population depends on the methods used, and each method seems to 
    have its own biases with respect to the way in which population history 
    affects parameter estimates. For instance, for \dadicli we see that estimates of 
    the shape parameter are most accurate in the YRI sample, while for \polydfe
    estimates are least accurate in the YRI sample. 
    With respect to the strength of selection, as in constant size populations,
    all methods are performing reasonably well, although \polydfe seems to
    lead in terms of accuracy.
\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things. That said, I think it would be good to plot the empirical (in the case simulated) SFS along with the SFSs from the MLE models. I would also do this for the demogrpahic models as well as the DFE models}

@@

    In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    from the same three DFE-inference packages on simulations of the
    vaquita porpoise genome. Importantly, these simulations use the vaquita demogrpahic model shown in Fig. X and a DFE that includes recessive deleterious mutations, as suggested in Robinsion et al. (2022). All methods
    perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter. This is not surprising as a mis-specified model is being fit to the data. All DFE inference methods used here assume an addive model (no dominance, $h = 1/2$) while the simulated data from the vaquita  was generated from a DFE in which more deleterious mutations are more recessive ($h < 1/2$)

    Reassuringly, for all three software packages, the inferred mean and shape of the distribution of selection coefficients 
    is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient
    and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
   \kelcomment{I think we should move this next sentence to the Discussion and expand on it a bit} Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
    so these results suggest caution in interpreting DFE inferences.

@@

    \subsection*{Simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. As stated above, simulations did or did not include a model 
    of (negative) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
    Three replicates of whole-genome simulations were performed for each species, with 
    a sample size of 100 diploid individuals. Demographic models are as stated above, each
    using the exon annotations from the human and vaquita porpoise genomes respectively.
    For simulations with selection, we sampled 10 diploid individuals from each of two populations (YRI and CEU). 
    To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations. \kelcomment{I'm confused about the 100 diploid individuals samples vs. the 10 diploid individuals sampled. Did some analyses use 100 and others use 10? That's fine, but maybe we can specify that more clearly...}

@@

    \subsection*{Distribution of fitness effects inference}
    We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
    \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
    Each software package was run using the same set of simulations.
    For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
    A key distinction between our analyses and those done in practice
    is that we have complete knowledge of which sites are under selection in the simulation,
    thus our inferences is a best case scenario. Empirical inference of the DFE is a much
    harder problem as some mutations assumed to be free of selection
    may in reality be under selection. As above, all methods were run using default settings.
    \kelcomment{I think we need a lot more detail here...how big was the coding sequence length? How was demogrpahic inference done for the different models via the different programs? How was the unscaling of E[s] done?, etc}
